---
title: "Hate Crime Trends"
author: "Candace Todd"
date: "December 17, 2019"
output: html_notebook
---


Become acquainted with the data
Explore intuition for the research questions before imposing formal structure (e.g., inferential or predictive dels)
Discover features in the data that influence modeling decisions

```{r, message=FALSE}
#Cleaning Environment
rm(list = ls())

#Loading necessary packages
library(openxlsx)
library(DataComputing)
library(tidyverse)
```

### Are Hate Crimes a Growing Problem in the U.S?
To answer this question, we'll look at Hate Crime Data and General Crime Data From the FBI, from the previous decade (excluding the current year).
  
  
  
#### Getting the Data
```{r, message=FALSE}
RawHate2018 <- openxlsx::read.xlsx(xlsxFile = "HateByState2018.xlsx")
RawHate2017 <- openxlsx::read.xlsx(xlsxFile = "HateByState2017.xlsx")
RawHate2016 <- openxlsx::read.xlsx(xlsxFile = "HateByState2016.xlsx")
RawHate2015 <- openxlsx::read.xlsx(xlsxFile = "HateByState2015.xlsx")
RawHate2014 <- openxlsx::read.xlsx(xlsxFile = "HateByState2014.xlsx")
RawHate2013 <- openxlsx::read.xlsx(xlsxFile = "HateByState2013.xlsx")
RawHate2012 <- openxlsx::read.xlsx(xlsxFile = "HateByState2012.xlsx")
RawHate2011 <- openxlsx::read.xlsx(xlsxFile = "HateByState2011.xlsx")
RawHate2010 <- openxlsx::read.xlsx(xlsxFile = "HateByState2010.xlsx")
RawHate2009 <- openxlsx::read.xlsx(xlsxFile = "HateByState2009.xlsx")
RawHate2008 <- openxlsx::read.xlsx(xlsxFile = "HateByState2008.xlsx")

RawCrime2018 <- openxlsx::read.xlsx(xlsxFile = "CrimeByState2018.xlsx")
RawCrime2017 <- openxlsx::read.xlsx(xlsxFile = "CrimeByState2017.xlsx")
RawCrime2016 <- openxlsx::read.xlsx(xlsxFile = "CrimeByState2016.xlsx")
RawCrime2015 <- openxlsx::read.xlsx(xlsxFile = "CrimeByState2015.xlsx")
RawCrime2014 <- openxlsx::read.xlsx(xlsxFile = "CrimeByState2014.xlsx")
RawCrime2013 <- openxlsx::read.xlsx(xlsxFile = "CrimeByState2013.xlsx")
RawCrime2012 <- openxlsx::read.xlsx(xlsxFile = "CrimeByState2012.xlsx")
RawCrime2011 <- openxlsx::read.xlsx(xlsxFile = "CrimeByState2011.xlsx")
RawCrime2010 <- openxlsx::read.xlsx(xlsxFile = "CrimeByState2010.xlsx")
RawCrime2009 <- openxlsx::read.xlsx(xlsxFile = "CrimeByState2009.xlsx")
RawCrime2008 <- openxlsx::read.xlsx(xlsxFile = "CrimeByState2008.xlsx")

```

  
Upon Inspection, we see that the FBI data is untidy, has awkward variable names, and has its variables classed in an unhelpful way. When downloading the data from [ucr.fbi.gov](https://ucr.fbi.gov/hate-crime/) I visually verified that the format of each General Crime Table and each Hate Crime Table respectively are almost identical throughout the years, so all tables have been inspected even though here I only preview one table.
```{r}
# Inspecting a Raw FBI General Crime Table
summary(RawCrime2008)
head(RawCrime2008, 15)

# Inspecting a Raw FBI Hate Crime Table
summary(RawHate2008)
head(RawHate2008, 12)
```
  
  
#### Tidying Up
  
Before anything can be done with the FBI data, we wrangle them into tidy format.
```{r}
# Cleaning the Raw FBI Tables
Crime2008 <- as.data.frame(RawCrime2008[-c(1:3),]) # removing header rows for general crime data
Crime2009 <- as.data.frame(RawCrime2009[-c(1:3),])
Crime2010 <- as.data.frame(RawCrime2010[-c(1:3),])
Crime2011 <- as.data.frame(RawCrime2011[-c(1:3),])
Crime2012 <- as.data.frame(RawCrime2012[-c(1:3),])
Crime2013 <- as.data.frame(RawCrime2013[-c(1:3),])
Crime2014 <- as.data.frame(RawCrime2014[-c(1:3),])
Crime2015 <- as.data.frame(RawCrime2015[-c(1:3),])
Crime2016 <- as.data.frame(RawCrime2016[-c(1:3),])
Crime2017 <- as.data.frame(RawCrime2017[-c(1:3),])
Crime2018 <- as.data.frame(RawCrime2018[-c(1:3),])

Hate2008 <- as.data.frame(RawHate2008[-c(1:3),]) # removing header rows for hate crime data
Hate2009 <- as.data.frame(RawHate2009[-c(1:3),])
Hate2010 <- as.data.frame(RawHate2010[-c(1:3),])
Hate2011 <- as.data.frame(RawHate2011[-c(1:4),])
Hate2012 <- as.data.frame(RawHate2012[-c(1:5),])
Hate2013 <- as.data.frame(RawHate2013[-c(1:5),])
Hate2014 <- as.data.frame(RawHate2014[-c(1:5),])
Hate2015 <- as.data.frame(RawHate2015[-c(1:5),])
Hate2016 <- as.data.frame(RawHate2016[-c(1:5),])
Hate2017 <- as.data.frame(RawHate2017[-c(1:5),])
Hate2018 <- as.data.frame(RawHate2018[-c(1:5),])

# Verifying
head(Crime2008, 5)
head(Hate2008, 5)
```

*Note: To avoid bulkiness of code, from here on out code chunks with repetitive actions that apply to the whole set of Crime or Hate tables in an identical manner (like above) will be suppressed. An example of the task being performed will be shown explicitly on one table from each category, but the code chunk with the same task being done on the other 11 tables will be hidden. However, all code will be visible in the .Rmd*  
*If all of the tables had been formatted identically, I would have used a for loop or an apply() function to wrangle multiple data sets, but because of slight variation in formatting from year to year I was more comfortable with this method of wrangling.*
  
Continuing the wrangling:
```{r}

rownames(Crime2008) <- 1:nrow(Crime2008) # re-numbering the rows
rownames(Hate2008) <- 1:nrow(Hate2008)

# Removing rows that appear to be footnotes
Crime2008 <- as.data.frame(Crime2008[-c(500:504),])
Hate2008 <- as.data.frame(Hate2008[-c(2388:2390),])

tail(Crime2008, 5)
tail(Hate2008, 5)

# Renaming the useful columns, re-classing variabels, and dropping (ignoring) columns that aren't of interest to the research question
Crime2008 <-
Crime2008 %>%
  transmute(state = tolower(gsub("(\\d*\\,*\\s*\\d+)$", "", `Table.5`)), # making all lowercase for consistency and removing numbers from states
            total = X2, # In the raw data, the `X2` column actually represents the area of the state where the crime occured,
            population = as.numeric(X4), # but we're only going to end up using the row with the state totals so we're renaming it `total` now.
            violentCrime = as.numeric(X5),#     All other crimes listed are a subset of violent or property crimes (according to the FBI website),
            propertyCrime = as.numeric(X10),#   so we'll only keep these two
            allCrime = (violentCrime + propertyCrime), # Creating variables I anticipate using in the future
            year = 2008
            )

Hate2008 <-
Hate2008 %>%
  transmute(state = tolower(gsub("(\\d*\\,*\\s*\\d+)$", "", `Table.13`)), 
            race_ethnicity = (as.numeric(X4) + as.numeric(X7)), # Race/Ethnicity/Ancestry is one column in later years so we combine them here
            religion = as.numeric(X5),
            sexualOrientation = as.numeric(X6),
            disability = as.numeric(X8),
            year = 2008
            )

str(Crime2008)
str(Hate2008)

# Fixing the Crime data so that there's a State name mext to the State total, because that's the row we want to preserve

# This for loop will grab the state name from the row above is there is no state name listed in the first column of the row
for (i in 1:nrow(Crime2008)){
  if(is.na(Crime2008[i,1])){
    Crime2008[i,1] <- Crime2008[(i-1),1]
  }
}
# Now the only rows without any NA values are the rows with the State Totals, which are the rows we want
Crime2008 <- na.omit(Crime2008)

# To filter the Hate data for the rows we want, we get rid of all rows who have an NA value in the first column
Hate2008 <-
  Hate2008 %>%
  mutate(Flag = ifelse(is.na(state), "drop", "keep")) %>% # assigning values to rows that give the state totals that I want to keep
  filter(Flag == "keep") %>%
  select(-c(Flag)) #removing my flag variable, as it has served its purpose

head(Crime2008)
head(Hate2008)
```
  
*This is where the block of code that cleans the 2009-2018 data sets is (in the .Rmd)*  
```{r, echo = FALSE, results = "hide"}
# Removing Footnotes
Crime2009 <- as.data.frame(Crime2009[-c(496:501),])
Crime2010 <- as.data.frame(Crime2010[-c(504:507),])
Crime2011 <- as.data.frame(Crime2011[-c(512:515),])
Crime2012 <- as.data.frame(Crime2012[-c(503:506),])
Crime2013 <- as.data.frame(Crime2013[-c(508:516),])
Crime2014 <- as.data.frame(Crime2014[-c(506:513),])
Crime2015 <- as.data.frame(Crime2015[-c(503:510),])
Crime2016 <- as.data.frame(Crime2016[-c(502:509),])
Crime2017 <- as.data.frame(Crime2017[-c(505:509),])
Crime2018 <- as.data.frame(Crime2018[-c(507:512),])

Hate2009 <- as.data.frame(Hate2009[-c(2275:2278),])
Hate2010 <- as.data.frame(Hate2010[-c(2188:2190),])
Hate2011 <- as.data.frame(Hate2011[-c(2184:2187),])
Hate2012 <- as.data.frame(Hate2012[-c(1957:1960),])
Hate2013 <- as.data.frame(Hate2013[-c(2071:2074),])
Hate2014 <- as.data.frame(Hate2014[-c(1896:1898),])
Hate2015 <- as.data.frame(Hate2015[-c(1981:1985),])
Hate2016 <- as.data.frame(Hate2016[-c(2017:2019),])
Hate2017 <- as.data.frame(Hate2017[-c(2298:2300),])
Hate2018 <- as.data.frame(Hate2018[-c(2283:2286),])

# Renaming the useful columns, re-classing variabels, and dropping (ignoring) columns that aren't of interest to the research question
Crime2009 <- Crime2009 %>%
  transmute(state = tolower(gsub("(\\d*\\,*\\s*\\d+)$", "", `Table.5`)), 
  total = X2, 
  population = as.numeric(X4), 
  violentCrime = as.numeric(X5),
  propertyCrime = as.numeric(X10),
  allCrime = (violentCrime + propertyCrime),
  year = 2009)

Crime2010 <- Crime2010 %>%
  transmute(state = tolower(gsub("(\\d*\\,*\\d+)$", "", `Table.5`)),
  total = X2, 
  population = as.numeric(X4), 
  violentCrime = as.numeric(X5),
  propertyCrime = as.numeric(X10),
  allCrime = (violentCrime + propertyCrime),
  year = 2010)

Crime2011 <- Crime2011 %>%
  transmute(state = tolower(gsub("(\\d*\\,*\\d+)$", "", `Table.5`)),
  total = X2, 
  population = as.numeric(X4), 
  violentCrime = as.numeric(X5),
  propertyCrime = as.numeric(X10),
  allCrime = (violentCrime + propertyCrime),
  year = 2011)

Crime2012 <- Crime2012 %>%
  transmute(state = tolower(gsub("(\\d*\\,*\\d+)$", "", `Table.5`)), 
  total = X2, 
  population = as.numeric(X4), 
  violentCrime = as.numeric(X5),
  propertyCrime = as.numeric(X10),
  allCrime = (violentCrime + propertyCrime),
  year = 2012)

Crime2013 <- Crime2013 %>%
  transmute(state = tolower(gsub("(\\d*\\,*\\d+)$", "", `Table.5`)),
  total = X2, 
  population = as.numeric(X4), 
  violentCrime = as.numeric(X5),
  propertyCrime = as.numeric(X11),
  allCrime = (violentCrime + propertyCrime),
  year = 2013)

Crime2014 <- Crime2014 %>%
  transmute(state = tolower(gsub("(\\d*\\,*\\d+)$", "", `Table.5`)),
  total = X2, 
  population = as.numeric(X4), 
  violentCrime = as.numeric(X5),
  propertyCrime = as.numeric(X11),
  allCrime = (violentCrime + propertyCrime),
  year = 2014)

Crime2015 <- Crime2015 %>%
  transmute(state = tolower(gsub("(\\d*\\,*\\d+)$", "", `Table.5`)),
  total = X2, 
  population = as.numeric(X4), 
  violentCrime = as.numeric(X5),
  propertyCrime = as.numeric(X11),
  allCrime = (violentCrime + propertyCrime),
  year = 2015)

Crime2016 <- Crime2016 %>%
  transmute(state = tolower(gsub("(\\d*\\,*\\d+)$", "", `Table.3`)), 
  total = X2, 
  population = as.numeric(X4), 
  violentCrime = as.numeric(X5),
  propertyCrime = as.numeric(X11),
  allCrime = (violentCrime + propertyCrime),
  year = 2016)

Crime2017 <- Crime2017 %>%
  transmute(state = tolower(gsub("(\\d*\\,*\\d+)$", "", `Table.5`)),
  total = X2, 
  population = as.numeric(X4), 
  violentCrime = as.numeric(X5),
  propertyCrime = as.numeric(X10),
  allCrime = (violentCrime + propertyCrime),
  year = 2017)

Crime2018 <- Crime2018 %>%
  transmute(state = tolower(gsub("(\\d*\\,*\\d+)$", "", `Table.5`)),
  total = X2, 
  population = as.numeric(X4), 
  violentCrime = as.numeric(X5),
  propertyCrime = as.numeric(X10),
  allCrime = (violentCrime + propertyCrime),
  year = 2018)

Hate2009 <- Hate2009 %>%
  transmute(state = tolower(gsub("(\\d*\\,*\\d+)$", "", `Table.13`)),
            race_ethnicity = (as.numeric(X4) + as.numeric(X7)),
            religion = as.numeric(X5),
            sexualOrientation = as.numeric(X6),
            disability = as.numeric(X8),
            year = 2009)

Hate2010 <- Hate2010 %>%
  transmute(state = tolower(gsub("(\\d*\\,*\\d+)$", "", `Table.13`)),
            race_ethnicity = (as.numeric(X4) + as.numeric(X7)),
            religion = as.numeric(X5),
            sexualOrientation = as.numeric(X6),
            disability = as.numeric(X8),
            year = 2010)

Hate2011 <- Hate2011 %>%
  transmute(state = tolower(gsub("(\\d*\\,*\\d+)$", "", `Table.13`)),
            race_ethnicity = (as.numeric(X4) + as.numeric(X7)),
            religion = as.numeric(X5),
            sexualOrientation = as.numeric(X6),
            disability = as.numeric(X8),
            year = 2011)

Hate2012 <- Hate2012 %>%
  transmute(state = tolower(gsub("(\\d*\\,*\\d+)$", "", `Table.13`)),
            race_ethnicity = (as.numeric(X4) + as.numeric(X7)),
            religion = as.numeric(X5),
            sexualOrientation = as.numeric(X6),
            disability = as.numeric(X8),
            year = 2012)

Hate2013 <- Hate2013 %>%
  transmute(state = tolower(gsub("(\\d*\\,*\\d+)$", "", `Table.13`)),
            race_ethnicity = (as.numeric(X4) + as.numeric(X7)),
            religion = as.numeric(X5),
            sexualOrientation = as.numeric(X6),
            disability = as.numeric(X8),
            gender = as.numeric(X9),
            genderIdentity = as.numeric(X10),
            year = 2013)

Hate2014 <- Hate2014 %>%
  transmute(state = tolower(gsub("(\\d*\\,*\\d+)$", "", `Table.13`)),
            race_ethnicity = (as.numeric(X4) + as.numeric(X7)),
            religion = as.numeric(X5),
            sexualOrientation = as.numeric(X6),
            disability = as.numeric(X8),
            gender = as.numeric(X9),
            genderIdentity = as.numeric(X10),
            year = 2014)

Hate2015 <- Hate2015 %>%
  transmute(state = tolower(gsub("(\\d*\\,*\\d+)$", "", `Table.13`)),
            race_ethnicity = as.numeric(X4),
            religion = as.numeric(X5),
            sexualOrientation = as.numeric(X6),
            disability = as.numeric(X7),
            gender = as.numeric(X8),
            genderIdentity = as.numeric(X9),
            year = 2015)

Hate2016 <- Hate2016 %>%
  transmute(state = tolower(gsub("(\\d*\\,*\\d+)$", "", `Table.13`)),
            race_ethnicity = as.numeric(X4),
            religion = as.numeric(X5),
            sexualOrientation = as.numeric(X6),
            disability = as.numeric(X7),
            gender = as.numeric(X8),
            genderIdentity = as.numeric(X9),
            year = 2016)

Hate2017 <- Hate2017 %>%
  transmute(state = tolower(gsub("(\\d*\\,*\\d+)$", "", `Table.13`)),
            race_ethnicity = as.numeric(X4),
            religion = as.numeric(X5),
            sexualOrientation = as.numeric(X6),
            disability = as.numeric(X7),
            gender = as.numeric(X8),
            genderIdentity = as.numeric(X9),
            year = 2017)

Hate2018 <- Hate2018 %>%
  transmute(state = tolower(gsub("(\\d*\\,*\\d+)$", "", `Table.13`)),
            race_ethnicity = as.numeric(X4),
            religion = as.numeric(X5),
            sexualOrientation = as.numeric(X6),
            disability = as.numeric(X7),
            gender = as.numeric(X8),
            genderIdentity = as.numeric(X9),
            year = 2018)

# Filtering the Crime and Hate data only for the rows we want to preserve

for (i in 1:nrow(Crime2009)){
  if(is.na(Crime2009[i,1])){
    Crime2009[i,1] <- Crime2009[(i-1),1]
  }
}
Crime2009 <- na.omit(Crime2009)

for (i in 1:nrow(Crime2010)){
  if(is.na(Crime2010[i,1])){
    Crime2010[i,1] <- Crime2010[(i-1),1]
  }
}
Crime2010 <- na.omit(Crime2010)

for (i in 1:nrow(Crime2011)){
  if(is.na(Crime2011[i,1])){
    Crime2011[i,1] <- Crime2011[(i-1),1]
  }
}
Crime2011 <- na.omit(Crime2011)

for (i in 1:nrow(Crime2012)){
  if(is.na(Crime2012[i,1])){
    Crime2012[i,1] <- Crime2012[(i-1),1]
  }
}
Crime2012 <- na.omit(Crime2012)

for (i in 1:nrow(Crime2013)){
  if(is.na(Crime2013[i,1])){
    Crime2013[i,1] <- Crime2013[(i-1),1]
  }
}
Crime2013 <- na.omit(Crime2013)

for (i in 1:nrow(Crime2014)){
  if(is.na(Crime2014[i,1])){
    Crime2014[i,1] <- Crime2014[(i-1),1]
  }
}
Crime2014 <- na.omit(Crime2014)

for (i in 1:nrow(Crime2015)){
  if(is.na(Crime2015[i,1])){
    Crime2015[i,1] <- Crime2015[(i-1),1]
  }
}
Crime2015 <- na.omit(Crime2015)

for (i in 1:nrow(Crime2016)){
  if(is.na(Crime2016[i,1])){
    Crime2016[i,1] <- Crime2016[(i-1),1]
  }
}
Crime2016 <- na.omit(Crime2016)

for (i in 1:nrow(Crime2017)){
  if(is.na(Crime2017[i,1])){
    Crime2017[i,1] <- Crime2017[(i-1),1]
  }
}
Crime2017 <- na.omit(Crime2017)

for (i in 1:nrow(Crime2018)){
  if(is.na(Crime2018[i,1])){
    Crime2018[i,1] <- Crime2018[(i-1),1]
  }
}
Crime2018 <- na.omit(Crime2018)

Hate2009 <- Hate2009 %>%
  mutate(Flag = ifelse(is.na(state), "drop", "keep")) %>%
  filter(Flag == "keep") %>%
  select(-c(Flag))

Hate2010 <- Hate2010 %>%
  mutate(Flag = ifelse(is.na(state), "drop", "keep")) %>%
  filter(Flag == "keep") %>%
  select(-c(Flag))

Hate2011 <- Hate2011 %>%
  mutate(Flag = ifelse(is.na(state), "drop", "keep")) %>%
  filter(Flag == "keep") %>%
  select(-c(Flag))

Hate2012 <- Hate2012 %>%
  mutate(Flag = ifelse(is.na(state), "drop", "keep")) %>%
  filter(Flag == "keep") %>%
  select(-c(Flag))

Hate2013 <- Hate2013 %>%
  mutate(Flag = ifelse(is.na(state), "drop", "keep")) %>%
  filter(Flag == "keep") %>%
  select(-c(Flag))

Hate2014 <- Hate2014 %>%
  mutate(Flag = ifelse(is.na(state), "drop", "keep")) %>%
  filter(Flag == "keep") %>%
  select(-c(Flag))

Hate2015 <- Hate2015 %>%
  mutate(Flag = ifelse(is.na(state), "drop", "keep")) %>%
  filter(Flag == "keep") %>%
  select(-c(Flag))

Hate2016 <- Hate2016 %>%
  mutate(Flag = ifelse(is.na(state), "drop", "keep")) %>%
  filter(Flag == "keep") %>%
  select(-c(Flag))

Hate2017 <- Hate2017 %>%
  mutate(Flag = ifelse(is.na(state), "drop", "keep")) %>%
  filter(Flag == "keep") %>%
  select(-c(Flag))

Hate2018 <- Hate2018 %>%
  mutate(Flag = ifelse(is.na(state), "drop", "keep")) %>%
  filter(Flag == "keep") %>%
  select(-c(Flag))
```
  
Some incomplete cleaning tasks that I couldn't debug slipped through the cracks, so here I employ "search and destroy" tacticts.
```{r}
Crime2008 <- Crime2008 %>%
  filter(state != "puerto rico") %>%
  mutate(state = gsub("2\\,\\s3", "", state))

Crime2009 <- Crime2009 %>%
  filter(state != "puerto rico") %>%
  mutate(state = gsub("2\\,\\s3", "", state))

Crime2013 <- Crime2013 %>%
  filter(state != "puerto rico") %>%
  mutate(state = gsub("4.", "", state))

Crime2014 <- Crime2014 %>%
  mutate(state = gsub("4\\,", "", state)) %>%
  filter(state != "puerto rico")

Crime2011 <- filter(Crime2011, state != "puerto rico")
Crime2012 <- filter(Crime2012, state != "puerto rico")
Crime2015 <- filter(Crime2015, state != "puerto rico")
Crime2016 <- filter(Crime2016, state != "puerto rico")
Crime2017 <- filter(Crime2017, state != "puerto rico")
Crime2018 <- filter(Crime2018, state != "puerto rico")

Hate2010 <- Hate2010[-51,]
```
  
Now we join all the `Crime` tables together and all the `Hate` tables together into longitudinal data frames.
```{r, message = FALSE}
# Joining all data
DecadeOfCrime <-
  full_join(Crime2008, Crime2009) %>%
  full_join(Crime2010) %>%
  full_join(Crime2011) %>%
  full_join(Crime2012) %>%
  full_join(Crime2013) %>%
  full_join(Crime2014) %>%
  full_join(Crime2015) %>%
  full_join(Crime2016) %>%
  full_join(Crime2017) %>%
  full_join(Crime2018) %>%
  select(-c(total))

DecadeOfHate <-
  full_join(Hate2008, Hate2009) %>%
  full_join(Hate2010) %>%
  full_join(Hate2011) %>%
  full_join(Hate2012) %>%
  full_join(Hate2013) %>%
  full_join(Hate2014) %>%
  full_join(Hate2015) %>%
  full_join(Hate2016) %>%
  full_join(Hate2017) %>%
  full_join(Hate2018) %>%
  replace_na(list(gender = 0, genderIdentity = 0))

head(DecadeOfCrime)
head(DecadeOfHate)
```

### EDA

### Analysis