---
title: "Economic Inequality and Hate Crimes"
output: html_notebook
---


Become acquainted with the data
Explore intuition for the research questions before imposing formal structure (e.g., inferential or predictive dels)
Discover features in the data that influence modeling decisions

```{r}
#Cleaning Environment
rm(list = ls())

#Loading necessary packages
library(openxlsx)
library(DataComputing)
library(tidyverse)
```

mention 538 and https://ucr.fbi.gov/hate-crime/2016/resource-pages/abouthatecrime

Explanations, definitions, and clarifications of variable names and the codes used for the variables were provided by the [read me in FiveThirtyEight's repo glossary page](https://ucr.fbi.gov/hate-crime/2016/resource-pages/abouthatecrime).

```{r}
#Importing the Data
CrimeMotiveState <- openxlsx::read.xlsx(xlsxFile = "Table_13_Hate_Crime_Incidents_per_Bias_Motivation_and_Quarter_by_State_and_Agency_2016.xlsx")
InequalityAndCrime538 <- readr::read_csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/hate-crimes/hate_crimes.csv")
PovertyByState <- readr::read_csv("State Poverty Levels 2016.txt")
ReadMe538 <- read_file("https://raw.githubusercontent.com/fivethirtyeight/data/master/hate-crimes/README.md")

```

```{r}
#Displaying the Data
PersonsExonerated



state.abb
state.area
state.center
state.division
state.name
state.region
state

```