---
title: "Economic Inequality and Hate Crimes"
output: html_notebook
---


Become acquainted with the data
Explore intuition for the research questions before imposing formal structure (e.g., inferential or predictive dels)
Discover features in the data that influence modeling decisions

```{r}
#Cleaning Environment
rm(list = ls())

#Loading necessary packages
library(openxlsx)
library(DataComputing)
library(tidyverse)
```

mention 538 and https://ucr.fbi.gov/hate-crime/2016/resource-pages/abouthatecrime

Explanations, definitions, and clarifications of variable names and the codes used for the variables were provided by the [read me in FiveThirtyEight's repo glossary page](https://ucr.fbi.gov/hate-crime/2016/resource-pages/abouthatecrime).

```{r}
#Importing the Data
IncomeInequality <- openxlsx::read.xlsx(xlsxFile = "publicspreadsheet.xlsx")
PovertyByState <-
CrimeTypeState <-

#Displaying the Data
PersonsExonerated



state.abb
state.area
state.center
state.division
state.name
state.region
state

```